`It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi

### 2023-04-06 manuscript/manuscript.qmd
Add authors, conclusions, acknowledgements

---
title: The invisible title
author:
  - name: Elizabeth Tilley
    orcid: 0000-0002-2095-9724
    email: tilleye@ethz.ch
    affiliations: 
        - id: ethz
          name: ETH Zurich, Global Health Engineering
          department: Department of Mechanical and Process Engineering
          address: Clausiusstrasse 37
          city: Zurich
          postal-code: 8092
    attributes:
        corresponding: true
  - name: Marc Kalina
    orcid: 0000-0002-6335-3845
    email: mkalina@ethz.ch
    affiliations:
        - ref: ethz
  - name: Jonathan Kwangulero
    orcid: 0000-0002-4050-5121
    email: kwangerlo@gmail.comm
    affilations:
        - id: polytech
          name: The Polytechnic, University of Malawi
          department: Department of Environmental Health
  - name: Hope Chilunga
    orcid: 0000-0003-4216-6319
    email: hchilunga@poly.ac.mw
    affilations:
        - ref: polytech
abstract: |
  Built in 1964, Malawi’s year of independence, Queen Elizabeth Hospital (QECH), or Queen’s, as it’s known locally, is the country’s largest public hospital. However, Queen’s is not served by regular municipal waste collection. What municipal collection that is done is ad-hoc, sporadic, and based on the hospital’s available financial resources. Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and openly burned, in several constantly smouldering piles, sending up clouds of smoke.  This study utilised a mixed-methods approach to both quantitatively measure the air quality around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers. Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3 air) were recorded every 5 minutes at 8 locations across the QECH for 2 months. Qualitative data collection consisted of 57 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).  Our results show that a) safe air quality thresholds are consistently exceeded across space and time; and b) that the most problematic air quality surrounds the shelter for caregivers and those receiving treatment for HIV/AIDs. Nearly every staff member interviewed recalled having received a complaint about the smoke, but those in charge of waste management felt powerless to make changes or deal with complaints.  Waste management interventions are desperately needed lest the patients who arrive at Queen’s leave with more health issues than the ones with which they arrived. 
keywords: 
  - waste management
  - healthcare waste
  - trash burning
  - air quality
  - Malawi
  - urbanisation
date: last-modified
bibliography: man-qechairquality.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Habitat International
      formatting: review
      model: 3p
      cite-style: authoryear
      highlights:
        - Low-cost monitors were used to assess air quality at Malawi’s largest hospital
        - Open waste burning is physically extensive and temporally frequent across the hospital campus
        - Air quality measured as PM2.5 and PM10 consistently exceeded even the most hazardous levels 
        - The most exposed locations were where guardians of patients stay and the HIV clinic
execute: 
  echo: false
  warning: false
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: console
---

@@

Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
(QECH), or Queen's, as it's known locally, is the country's largest
public hospital. Designed in the Tropical Modernism style of
architecture popular within Africa in the late colonial period, Queen's
open-plan form sprawls across a broad swath of central Blantyre, with
dozens of wards, specialised facilities, and administrative buildings,
linked together by a bewildering maze of covered walkways. A hive of
around-the-clock activity, Queen's bustles from dawn to dusk with a
constant stream of patients, drawn from across the country, and
maintains a city-like feel even after dark as staff and stay-in family
members cook and care for their in-patient dependants. Yet, for the
first-time visitor, it's not the architecture or the crowds that are the
most striking when entering the grounds, but the smell. Immediately
noticeable, even from the road outside the hospital, Queen's is
smothered by an acrid-smelling, white smoke, which hangs over the
grounds day and night: the product of several constantly burning fires
spread across the campus. Breathing this air, when a significant amount
of burning is occurring, primarily at night when the smoke is less
visible, can be incredibly uncomfortable. Moreover, the tropical
architecture, designed to let air and light circulate, means that even
indoors, within patient wards, surgery, and treatment areas, the air
quality can also be unbearable. Poor air quality associated with
burning, through released particulate matter (PM), has been linked to
multiple negative health outcomes like asthma, heart disease, heart
failure, stroke, and cancer among others [@anderson2012clearing;
@bell2012assessment]. In a space of healing and recovery, why is such
burning occurring, and how can it continue to persist?

@@

>>>>

---
title: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"
author:
  - name: Elizabeth Tilley
    orcid: 0000-0002-2095-9724
    email: tilleye@ethz.ch
    affiliations: 
        - id: ethz
          name: ETH Zurich
          department: Department of Mechanical and Process Engineering
          city: Zurich
          country: Switzerland
    attributes:
        corresponding: true
  - name: Hope Chilunga
    orcid: 0000-0003-4216-6319
    email: hchilunga@poly.ac.mw
    affiliations: 
        - id: mubas
          name: Malawi University of Business and Applied Sciences
          department: Department of Environmental Health
          city: Blantyre
          country: Malawi
  - name: Jonathan Kwangulero
    orcid: 0000-0002-4050-5121
    email: kwangerlo@gmail.comm
    affilations:
        - ref: mubas
  - name: Lars Schöbitz
    orcid: 0000-0003-2196-5015
    email: lschoebitz@ethz.ch
    affiliations:
        - ref: ethz
  - name: Heiko Heilgendorff
    email: me@me.com
    affiliations: 
        - id: ukzn
          name: University of KwaZulu-Natal
          department: School of Mathematics, Statistics and Computer Science
          city: Durban
          country: South Africa
  - name: Marc Kalina
    orcid: 0000-0002-6335-3845
    email: mkalina@ethz.ch
    affiliations:
        - ref: ethz
abstract: |
  Blantyre, Malawi’s Queen Elizabeth Central Hospital (QECH), or Queen’s, as it’s known locally, is the country’s largest public hospital. However, Queen’s is not served by regular municipal waste collection. What municipal collection that is done is ad-hoc, sporadic, and based on the hospital’s available financial resources. Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and openly burned, in several constantly smouldering piles, sending up clouds of smoke.  Speaking directly to an identified knowledge gap on air quality impacts linked to trash burning and the paucity of African urban dwellers’ voices on air quality issues, this study employed a mixed-methods approach to both quantitatively measure the air quality around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers. Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3 air) were recorded every 5 minutes at 8 locations across the QECH for 2 months. Qualitative data collection consisted of 57 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).  Our results show that safe air quality thresholds are consistently exceeded across space and time and that the most problematic air quality surrounds the shelter for caregivers and those receiving treatment for HIV/AIDs. Moreover, staff and visitors are severely impacted by the poor air quality within the space, but  feel powerless to make changes or address complaints. Waste management interventions are desperately needed lest the patients who arrive at Queen’s leave with more health issues than the ones with which they arrived. 
keywords: 
  - waste management
  - healthcare waste
  - trash burning
  - air quality
  - Malawi
  - urbanisation
date: last-modified
bibliography: man-qechairquality.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Habitat International
      formatting: review
      model: 3p
      cite-style: authoryear
      highlights:
        - Low-cost monitors were used to assess air quality at Malawi’s largest hospital
        - Open waste burning is physically extensive and temporally frequent across the hospital campus
        - Air quality measured as PM2.5 and PM10 consistently exceeded even the most hazardous levels 
        - The most exposed locations were where guardians of patients stay and the HIV clinic
execute: 
  echo: false
  warning: false
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: console
---

@@

Built in 1964, Malawi's year of independence, Queen Elizabeth Central Hospital
(QECH), or Queen's, as it's known locally, is the country's largest
public hospital. Designed in the Tropical Modernism style of
architecture popular within Africa in the late colonial period, Queen's
open-plan form sprawls across a broad swath of central Blantyre, with
dozens of wards, specialised facilities, and administrative buildings,
linked together by a bewildering maze of covered walkways. A hive of
around-the-clock activity, Queen's bustles from dawn to dusk with a
constant stream of patients, drawn from across the country, and
maintains a city-like feel even after dark as staff and stay-in family
members cook and care for their in-patient dependants. Yet, for the
first-time visitor, it's not the architecture or the crowds that are the
most striking when entering the grounds, but the smell. Immediately
noticeable, even from the road outside the hospital, Queen's is
smothered by an acrid-smelling, white smoke, which hangs over the
grounds day and night: the product of several constantly burning fires
spread across the campus. Breathing this air, when a significant amount
of burning is occurring, primarily at night when the smoke is less
visible, can be incredibly uncomfortable. Moreover, the tropical
architecture, designed to let air and light circulate, means that even
indoors, within patient wards, surgery, and treatment areas, the air
quality can also be unbearable. Poor air quality associated with
burning, through released particulate matter (PM), has been linked to
multiple negative health outcomes like asthma, heart disease, heart
failure, stroke, and cancer among others [@anderson2012clearing;
@bell2012assessment]. In a space of healing and recovery, why is such
burning occurring, and how can it continue to persist?

@@

The air quality, as a result of open waste burning at Queen’s is poor and not suitable for a city, let alone a hospital. Over the course of this two-month study, the hazardous limits for both PM parameters were exceeded at all locations except for the Lions Sight hospital. The WHO limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the most hazardous limit. The composition of emissions recorded ranged significantly, analysed as the ratio PM10:PM2.5 likely reflects a range of burning styles and/or fuel (waste) type, and no discernable trend was observed..  Among interview respondents, there was a general  consensus that air quality caused by the burning of waste within the hospital was a problem, although there were significant differences between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials. For most, going home or leaving the Queen’s campus was the only pathway to relief, though for patients who can not leave, endurance was the only option. This work was originally planned as a before-and-after study:  we envisioned reporting these baseline results compared to follow-up measurements once the long-promised incinerator was commissioned.  However, nearly 4 years later, the incinerator is rarely functional (partly for technical reasons, partly for financial ones) and the smoke hovering over Queen’s is still thick.  Waste management has long been known as an environmental health issue, but in the case of open trash burning at a medical facility, it becomes more:  for some of the most vulnerable, especially those with TB, asthma or other chronic disease, it could be a matter of life or death.

# Acknowledgements

The authors are grateful to the staff, patients and caregivers at QECH for their assistance, willingness to share their views, and importantly, perseverance in delivering the best care possible under difficult circumstances. Moreover, we would like to acknowledge the assistance and effort of Open Data Durban, now the Open Cities Lab, for their assistance in constructing and programming the sensors.

### 2023-04-06 manuscript/manuscript.tex
Add authors, conclusions, acknowledgements

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Habitat International}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The invisible title},
  pdfauthor={Elizabeth Tilley; Marc Kalina; Jonathan Kwangulero; Hope Chilunga},
  pdfkeywords={waste management, healthcare waste, trash burning, air
quality, Malawi, urbanisation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{The invisible title}
\author[1]{Elizabeth Tilley%
\corref{cor1}%
}
 \ead{tilleye@ethz.ch} 
\author[1]{Marc Kalina%
%
}
 \ead{mkalina@ethz.ch} 
\author[]{Jonathan Kwangulero%
%
}
 \ead{kwangerlo@gmail.comm} 
\author[]{Hope Chilunga%
%
}
 \ead{hchilunga@poly.ac.mw}

\affiliation[1]{organization={ETH Zurich, Global Health
Engineering, Department of Mechanical and Process
Engineering},addressline={Clausiusstrasse
37},city={Zurich},postcode={8092},postcodesep={}}

@@

\begin{abstract}
Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
(QECH), or Queen's, as it's known locally, is the country's largest
public hospital. However, Queen's is not served by regular municipal
waste collection. What municipal collection that is done is ad-hoc,
sporadic, and based on the hospital's available financial resources.
Rather, most hospital waste (infectious and non-infectious) is gathered
by grounds staff and openly burned, in several constantly smouldering
piles, sending up clouds of smoke. This study utilised a mixed-methods
approach to both quantitatively measure the air quality around QECH, and
to qualitatively investigate the perceived impacts amongst staff and
caregivers. Low-cost sensors measuring particulate matter (PM) with
particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5),
expressed as the mass of PM per volume of air (μg PMx/m3 air) were
recorded every 5 minutes at 8 locations across the QECH for 2 months.
Qualitative data collection consisted of 57 interviews with patients,
caregivers and hospital staff (including janitorial and maintenance
staff, nurses, doctors, and administrators). Our results show that a)
safe air quality thresholds are consistently exceeded across space and
time; and b) that the most problematic air quality surrounds the shelter
for caregivers and those receiving treatment for HIV/AIDs. Nearly every
staff member interviewed recalled having received a complaint about the
smoke, but those in charge of waste management felt powerless to make
changes or deal with complaints. Waste management interventions are
desperately needed lest the patients who arrive at Queen's leave with
more health issues than the ones with which they arrived.
\end{abstract}

@@

\begin{keyword}
    waste management \sep healthcare waste \sep trash burning \sep air
quality \sep Malawi \sep 
    urbanisation
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable]}{\end{tcolorbox}}\fi

@@

>>>>

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Habitat International}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={`It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi},
  pdfauthor={Elizabeth Tilley; Hope Chilunga; Jonathan Kwangulero; Lars Schöbitz; Heiko Heilgendorff; Marc Kalina},
  pdfkeywords={waste management, healthcare waste, trash burning, air
quality, Malawi, urbanisation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{`It Is Unbearable To Breath Here': Air Quality, Open
Incineration, And Misinformation In Blantyre, Malawi}
\author[1]{Elizabeth Tilley%
\corref{cor1}%
}
 \ead{tilleye@ethz.ch} 
\author[2]{Hope Chilunga%
%
}
 \ead{hchilunga@poly.ac.mw} 
\author[]{Jonathan Kwangulero%
%
}
 \ead{kwangerlo@gmail.comm} 
\author[1]{Lars Schöbitz%
%
}
 \ead{lschoebitz@ethz.ch} 
\author[3]{Heiko Heilgendorff%
%
}
 \ead{me@me.com} 
\author[1]{Marc Kalina%
%
}
 \ead{mkalina@ethz.ch}

\affiliation[1]{organization={ETH Zurich, Department of Mechanical and
Process
Engineering},city={Zurich},country={Switzerland},countrysep={,},postcodesep={}}
\affiliation[2]{organization={Malawi University of Business and Applied
Sciences, Department of Environmental
Health},city={Blantyre},country={Malawi},countrysep={,},postcodesep={}}
\affiliation[3]{organization={University of KwaZulu-Natal, School of
Mathematics, Statistics and Computer
Science},city={Durban},country={South
Africa},countrysep={,},postcodesep={}}

@@

\begin{abstract}
Blantyre, Malawi's Queen Elizabeth Central Hospital (QECH), or Queen's,
as it's known locally, is the country's largest public hospital.
However, Queen's is not served by regular municipal waste collection.
What municipal collection that is done is ad-hoc, sporadic, and based on
the hospital's available financial resources. Rather, most hospital
waste (infectious and non-infectious) is gathered by grounds staff and
openly burned, in several constantly smouldering piles, sending up
clouds of smoke. Speaking directly to an identified knowledge gap on air
quality impacts linked to trash burning and the paucity of African urban
dwellers' voices on air quality issues, this study employed a
mixed-methods approach to both quantitatively measure the air quality
around QECH, and to qualitatively investigate the perceived impacts
amongst staff and caregivers. Low-cost sensors measuring particulate
matter (PM) with particle sizes less than 10μm (PM10) and less than
2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3
air) were recorded every 5 minutes at 8 locations across the QECH for 2
months. Qualitative data collection consisted of 57 interviews with
patients, caregivers and hospital staff (including janitorial and
maintenance staff, nurses, doctors, and administrators). Our results
show that safe air quality thresholds are consistently exceeded across
space and time and that the most problematic air quality surrounds the
shelter for caregivers and those receiving treatment for HIV/AIDs.
Moreover, staff and visitors are severely impacted by the poor air
quality within the space, but feel powerless to make changes or address
complaints. Waste management interventions are desperately needed lest
the patients who arrive at Queen's leave with more health issues than
the ones with which they arrived.
\end{abstract}

@@

\begin{keyword}
    waste management \sep healthcare waste \sep trash burning \sep air
quality \sep Malawi \sep 
    urbanisation
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

@@

The air quality, as a result of open waste burning at Queen's is poor
and not suitable for a city, let alone a hospital. Over the course of
this two-month study, the hazardous limits for both PM parameters were
exceeded at all locations except for the Lions Sight hospital. The WHO
limits were exceeded fewer than 50 times at five of the locations, and
only the monitors at the Lighthouse Clinic and the Guardian Shelter
recorded more than 50 instances above the most hazardous limit. The
composition of emissions recorded ranged significantly, analysed as the
ratio PM10:PM2.5 likely reflects a range of burning styles and/or fuel
(waste) type, and no discernable trend was observed.. Among interview
respondents, there was a general consensus that air quality caused by
the burning of waste within the hospital was a problem, although there
were significant differences between respondents over their
understandings of potential impacts, the effectiveness of various coping
mechanisms, and their problematisations linked to the burning of
specific waste materials. For most, going home or leaving the Queen's
campus was the only pathway to relief, though for patients who can not
leave, endurance was the only option. This work was originally planned
as a before-and-after study: we envisioned reporting these baseline
results compared to follow-up measurements once the long-promised
incinerator was commissioned. However, nearly 4 years later, the
incinerator is rarely functional (partly for technical reasons, partly
for financial ones) and the smoke hovering over Queen's is still thick.
Waste management has long been known as an environmental health issue,
but in the case of open trash burning at a medical facility, it becomes
more: for some of the most vulnerable, especially those with TB, asthma
or other chronic disease, it could be a matter of life or death.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

The authors are grateful to the staff, patients and caregivers at QECH
for their assistance, willingness to share their views, and importantly,
perseverance in delivering the best care possible under difficult
circumstances. Moreover, we would like to acknowledge the assistance and
effort of Open Data Durban, now the Open Cities Lab, for their
assistance in constructing and programming the sensors.
